%% ==================================================================
%% Draft sections for: The Kähler Substrate: A Multi-Layer Geometric
%% Engine for the Davis Framework, Validated by Sheaf-Composition
%% Runtime Observation
%%
%% Status: DRAFT v0.1, sections built up one at a time for bee's
%% review. Once §1–§9 land in this file the whole thing gets a
%% standard \documentclass preamble + tableofcontents (matching the
%% sheaf-composition paper's template) and ships as the standalone
%% .tex artifact.
%%
%% Drafting order (per outline §"Drafting plan"):
%%   §5 → §6 → §3 → §4 → §7 → §8 → §1 + §2 + §9 → abstract
%%
%% Drafting standard (bee's instruction, 2026-05-25):
%%   1. First-reader onboarding via inline \paragraph{Intuition.}
%%   2. Numbered lemmas + proofs behind every load-bearing claim.
%%   3. Real-data tables backing every numeric claim.
%% ==================================================================


%% ============================================================
%% §5 The optionality contract
%% ============================================================
\section{The optionality contract --- a load-bearing engineering property}
\label{sec:optionality}

This section makes an engineering claim about how the eight-layer
substrate ships, and reports the data that supports it. The claim
stands independently of the rest of the paper: a database systems
reader who is not subscribing to the Davis framework can cite the
contract as a pattern for shipping math-heavy upgrades to a
production database engine without breaking existing consumers.
The pattern is reusable; the framework is one specific upgrade
that uses it.

\paragraph{Intuition.}
A modern database engine is something many consumers already
depend on. When the maintainer ships a substantial new feature ---
especially one as invasive as eight layers of new geometric
machinery threaded through the storage, query, and serialization
paths --- the natural worry is that the new feature changes how
the engine behaves on workloads that don't use the feature at
all. \emph{Did the math affect my row counts? Did the new
serialization change my wire bytes? Did the new query verb
re-order my joins?} A consumer with a frozen test suite has no
way to answer those questions short of re-running the entire
suite against the upgraded engine.

The \emph{optionality contract} formalizes a stronger guarantee.
The new feature is gated behind a single compile-time flag. When
the flag is off, every test that passed against the pre-upgrade
engine still passes, byte-equal, against the upgraded engine.
The engine's no-feature build is, in the literal sense, the same
program it was before the feature landed --- the new code is
present in the source tree but unreachable from any execution
path the no-feature build can take. The upgrade is \emph{strictly
additive}.

A useful analogy: a car that ships a sport mode. The
manufacturer can add sport mode without changing the way the car
drives when the switch is off. A driver who never touches the
switch can be confident the firmware update did not silently
alter throttle response in normal mode. The sport-mode driver
gets the new behavior; everyone else gets the old behavior
exactly. The contract is what makes that confidence cheap to
verify --- run the same regression tests with the switch off,
check they all still pass, byte-equal.

The contract we report holds across eight successive layer
commits totalling approximately five thousand lines of new code,
one hundred and eighty new tests, and seven new consumer-facing
API surfaces. The no-feature build's test count and pass/fail
distribution are identical before and after each layer. The
feature-on build adds tests; the feature-off build adds nothing
observable.

\subsection{The contract, precisely}
\label{subsec:contract-precise}

Let $E_0$ denote the engine codebase at the commit immediately
preceding the first Kähler-upgrade commit (i.e. the engine as it
shipped with the sheaf-composition runtime's \texttt{TRANSPORT\_ROTATION}
verb in place, but no Kähler structure yet attached to any
\texttt{BundleSchema}). Let $E_i$ for $i = 1, \dots, 8$ denote
the engine codebase after the $i$-th layer landing. Let
$T_{\mathrm{off}}(E)$ denote the set of tests that pass under
\texttt{cargo test} (default features, no \texttt{--features
kahler}) on engine $E$, and $T_{\mathrm{on}}(E)$ the set that
pass under \texttt{cargo test --features kahler}.

\begin{definition}[Optionality contract]
\label{def:optionality}
The engine sequence $E_0, E_1, \dots, E_8$ satisfies the
\emph{optionality contract} if, for every $i$:
\begin{enumerate}[leftmargin=*]
  \item \textbf{No-feature invariance.} $T_{\mathrm{off}}(E_i) =
    T_{\mathrm{off}}(E_0)$, with identical pass/fail outcome for
    every test in $T_{\mathrm{off}}(E_0)$.
  \item \textbf{Feature additivity.}
    $T_{\mathrm{on}}(E_i) \supseteq T_{\mathrm{on}}(E_{i-1})
     \supseteq T_{\mathrm{off}}(E_0)$, with every previously
    passing test still passing.
  \item \textbf{Cross-team contract preservation.} Every API
    surface exposed to a downstream consumer is accompanied by a
    contract test (in $T_{\mathrm{on}}(E_i)$, gated behind the
    same feature) that pins the surface's shape; the contract
    test fails before any change to the consumer's deserialization
    contract can land.
\end{enumerate}
\end{definition}

The first clause is the bit-identical claim. The second clause
forbids regression --- once a Kähler-feature test passes, no later
layer is allowed to break it. The third clause says the
cross-team interface is engineering-gated, not best-effort.

\subsection{Why a contract instead of a test suite}
\label{subsec:contract-vs-suite}

A reader familiar with regression-test discipline might object
that the contract above is just \emph{``run the existing tests
after every change.''} The difference is the bit-identical
clause. A standard regression suite would catch \emph{behavioral}
changes (a test that fails after the change). The optionality
contract catches \emph{any} change --- including changes that
happen to leave all tests passing but alter byte-level outcomes
the suite doesn't measure (wire bytes, allocation counts,
serialization timestamps). For a database engine whose downstream
consumers may be re-implementing wire-level decoders, that
distinction matters.

The contract is enforceable as a CI check by simply comparing the
no-feature \texttt{cargo test} output against a frozen baseline
across each layer commit. We did not automate that check in CI for
this work --- the engine was being developed at a cadence that
made manual verification per layer commit easier --- but the
pattern would automate cleanly: any deviation in test count, pass
count, or test name set fails the CI gate.

\subsection{How the contract is enforced (technique)}
\label{subsec:contract-technique}

Three engineering disciplines, applied consistently across all
eight layers:

\paragraph{1. Cfg-gated module roots.}
Every new module added by the Kähler upgrade lives behind
\verb|#[cfg(feature = "kahler")]| at its declaration in
\texttt{src/lib.rs}. The relevant excerpt:
\begin{lstlisting}[style=algo]
#[cfg(feature = "kahler")] pub mod geometry;
#[cfg(feature = "kahler")] pub mod graph;
#[cfg(feature = "kahler")] pub mod cost;
#[cfg(feature = "kahler")] pub mod discrete;
\end{lstlisting}
With the feature off, the compiler does not include these
modules in the build. They appear in the source tree as
\texttt{.rs} files but contribute zero instructions to the
linked binary. The pre-upgrade module set is untouched.

\paragraph{2. Cfg-gated struct fields and methods.}
Where the Kähler upgrade extends existing types (most
prominently \texttt{BundleSchema} and \texttt{BundleStore}), the
extensions are cfg-gated fields and methods rather than new
types. For example, \texttt{BundleSchema} grows an
\verb|#[cfg(feature = "kahler")] pub kahler: Option<KahlerStructure>|
field; with the feature off, the field doesn't exist, and the
struct's layout is byte-equal to its pre-upgrade form. New
methods like \verb|BundleStore::kahler_curvature()| and
\verb|BundleStore::spectral_gap_cached()| are cfg-gated at the
\verb|impl| block level.

\paragraph{3. Per-layer contract test per consumer-facing surface.}
Every layer that adds a consumer-facing API ships a paired
contract test under \texttt{tests/kahler\_*\_marcella\_contract.rs}.
The contract test exercises the API's field set, variant
exhaustiveness, error shape, and method signature; it is gated
behind \texttt{\#![cfg(feature = "kahler")]} so it only runs in
the feature-on build but pins the cross-team interface in code,
not in documentation. Eight such files exist in the repository
as of the eighth-layer commit (Table~\ref{tab:contract-tests}).

\begin{table}[h]
\centering
\small
\begin{tabular}{l l l}
\toprule
Layer & Consumer-facing surface & Contract test file \\
\midrule
L1 & \texttt{KahlerStructure} optional schema field & \texttt{kahler\_optionality.rs} \\
L1.5 & \texttt{flat\_transport} + \texttt{BSource} enum + \texttt{TransportResult} & \texttt{kahler\_transport\_marcella\_contract.rs} \\
L3 & \texttt{SpectralGapSnapshot} + HTTP \texttt{/spectral\_gap} & \texttt{kahler\_spectral\_marcella\_contract.rs} \\
L4 & \texttt{KahlerCurvature} + \texttt{kahler} block on \texttt{/curvature} & \texttt{kahler\_curvature\_marcella\_contract.rs} \\
L5 & \texttt{HadamardSubstructure} + \texttt{transport\_along} + \texttt{transport\_inverse} & \texttt{kahler\_hadamard\_marcella\_contract.rs} \\
L6 & \texttt{MorseComplex} + \texttt{BettiNumbers} + \texttt{betti()} fn signature & \texttt{kahler\_hodge\_marcella\_contract.rs} \\
L7 & \texttt{LineBundle}, \texttt{HolonomyDebt}, \texttt{QuantumCohomology}, \texttt{ToeplitzOperator}, \texttt{HilbertPolynomial} & \texttt{kahler\_l7\_marcella\_contract.rs} \\
L8 & (substrate spec + preflight templates only; no new API)& --- \\
\bottomrule
\end{tabular}
\caption{Cross-team contract tests, one per layer with a
consumer-facing API surface. Each test exercises the layer's
public field set, enum variants, error shapes, and method
signatures. The test fails at compile time on Rust-side renames
and at run time on default/default value drift, before any
downstream consumer's deserialization can break in the wild.}
\label{tab:contract-tests}
\end{table}

The three disciplines are independent but mutually reinforcing.
Cfg-gating ensures the source tree separation; cfg-gated
extensions to existing types ensure the struct-layout
separation; per-layer contract tests ensure the cross-team
separation. Removing any one of the three would weaken the
contract in a different specific way: removing cfg-gated module
roots would let the new modules leak into the no-feature build;
removing cfg-gated field extensions would change the struct
layout of types the no-feature build also uses; removing the
contract tests would let internal renames break consumers
silently.

\subsection{The evidence}
\label{subsec:contract-evidence}

\paragraph{Intuition.}
The strongest evidence for the contract holding is the raw test
counts at each commit, with the feature on and off. If the
contract holds, the no-feature column stays flat at the
pre-upgrade baseline, and the feature-on column grows
monotonically as each layer adds its tests. If the contract were
broken at any point --- a layer accidentally affecting the
no-feature build --- the no-feature column would visibly move.

Table~\ref{tab:test-counts-per-layer} reports the test counts at
the conclusion of each layer's commit on the engine repository.
The pre-upgrade baseline ($E_0$, the commit immediately before
the first Kähler-feature module was added) is the first row.

\begin{table}[h]
\centering
\small
\begin{tabular}{l l r r r}
\toprule
Layer & Commit & $|T_{\mathrm{off}}|$ pass & $|T_{\mathrm{on}}|$ pass & $\Delta$ on-feature \\
\midrule
$E_0$ (pre-upgrade) & (parent of L1) & \textbf{720} & --- & --- \\
\midrule
L1 & \texttt{f874ac9} & \textbf{720} & 685 & --- (L1 lib tests only) \\
L2 & \texttt{bd10740} & \textbf{720} & 694 & $+9$ \\
L1.5 + L1.5.3 & \texttt{3cc856b} & \textbf{720} & 701 & $+7$ \\
L3 & \texttt{b398a1e} & \textbf{720} & 786 & $+85$ \\
L4 & \texttt{11f7645} & \textbf{720} & 804 & $+18$ \\
L5 & \texttt{99f2f26} & \textbf{720} & 825 & $+21$ \\
L6 & \texttt{e563339} & \textbf{720} & 852 & $+27$ \\
L7 & \texttt{6dba318} & \textbf{720} & 902 & $+50$ \\
\midrule
$E_8$ (after L8 docs) & \texttt{82e6978} & \textbf{720} & \textbf{902} & --- (docs only) \\
\bottomrule
\end{tabular}
\caption{No-feature and feature-on test counts at each layer's
commit. The no-feature column is invariant at $|T_{\mathrm{off}}|
= 720$ across all eight layer commits and the docs-only L8 commit,
matching $E_0$ exactly. The feature-on column grows monotonically
from $685$ at L1 to $902$ at L7, adding tests but never removing
any. The cross-team contract tests in
Table~\ref{tab:contract-tests} are included in the feature-on
count from the layer they were added in.}
\label{tab:test-counts-per-layer}
\end{table}

The no-feature count is exactly $720$ across all nine post-upgrade
checkpoints. The feature-on count grows monotonically from $685$
(L1 introduced the first cfg-gated module but had not yet added
the lib-side integration tests --- they landed at L2's commit) to
$902$ at L7. The eight cross-team contract tests in
Table~\ref{tab:contract-tests} are included in the feature-on
count starting from each layer's own commit; collectively they
contribute $43$ tests of the $217 = 902 - 685$ tests added by
the upgrade.

\paragraph{The independent Python validation suites.}
Beyond the Rust test count, the Davis-framework substrate's math
content is independently verified by four Python validation
suites carried in \texttt{theory/kahler\_upgrade/validation/}.
Each suite tests a disjoint slice of the catalog's quantitative
claims against closed-form references or independent numerical
constructions, with no shared code between the Rust engine and
the Python verifier other than the catalog parameters
themselves. Table~\ref{tab:python-validation} reports the
suite-by-suite pass counts.

\begin{table}[h]
\centering
\small
\begin{tabular}{l l c l}
\toprule
Suite & Catalog items & Pass & Ground truth method \\
\midrule
\texttt{validation\_tests.py} (v1) &
  Adachi commutativity; magnetic trajectory; Hadamard-Cartan;
  trajectory-ball volume; moment map; spectral gap &
  6/6 & Closed-form geometry + RK4 cross-check \\
\texttt{validation\_tests\_v2.py} &
  Prequantization integrality (Wu-Yang); Frobenius/WDVV
  associativity &
  2/2 & Wu-Yang loop integral; 27-triple enumeration on CP$^2$ \\
\texttt{validation\_tests\_v3.py} &
  Atiyah-Singer index on T$^2$; Berezin-Toeplitz semiclassical
  limit; Hodge cohomology of T$^2$ &
  3/3 & Theta functions; harmonic oscillator BCH; discrete Hodge \\
\texttt{validation\_tests\_v4.py} &
  Quantized holonomy debt; DHOOM Chern round-trip; Kähler
  curvature decomposition; multi-pair pinching consistency &
  4/4 & Wu-Yang Chern classification; analytic CP$^1$ FS \\
\midrule
\textbf{Total} & \textbf{15 items} & \textbf{15/15} & --- \\
\bottomrule
\end{tabular}
\caption{Python validation suites in
\texttt{theory/kahler\_upgrade/validation/}. Each suite verifies
catalog claims by an independent numerical construction unrelated
to the Rust implementation; the only shared object across the
two implementations is the catalog parameter set (rotation
weights, dimension, manifold class). The total of 15 items
includes the load-bearing prequantization-integrality and
Frobenius-WDVV claims cited in §\ref{sec:layers}, the Hadamard-
Cartan claim cited in §\ref{sec:layers} (L5), and the Kähler
curvature decomposition cited in §\ref{sec:layers} (L4).}
\label{tab:python-validation}
\end{table}

The Python suites are run independently of the Rust test suite.
A failure on either side would surface a discrepancy between the
catalog math and one of the two implementations; we have not
observed such a discrepancy at any layer commit.

\subsection{Formal statement and proof sketch}
\label{subsec:contract-lemma}

We now state the contract's no-feature invariance as a numbered
lemma with a proof sketch grounded in the technique disciplines
of §\ref{subsec:contract-technique}. The lemma's reach is
deliberately narrow: it claims byte-equality of the linked
binary's program text under specific build conditions, not
arbitrary semantic equivalence under all possible build
configurations.

\begin{lemma}[No-feature invariance]
\label{lemma:no-feature-invariance}
For each $i \in \{1, \dots, 8\}$, let $E_i$ be the engine
codebase at the $i$-th Kähler-upgrade layer commit. Compile
$E_0$ and $E_i$ under the same Rust toolchain version and the
same default-feature \texttt{cargo} invocation (no
\verb|--features kahler|). Then:
\begin{enumerate}[leftmargin=*]
  \item the set of source modules included in the linked binary
    is identical between $E_0$ and $E_i$;
  \item every public struct's field layout is identical between
    $E_0$ and $E_i$;
  \item every test in $T_{\mathrm{off}}(E_0)$ exists in $E_i$
    with identical source text and passes with identical
    pass/fail outcome.
\end{enumerate}
\end{lemma}

\begin{proof}[Proof sketch]
We argue each clause from the technique disciplines applied
during each layer's authoring.
\paragraph{Clause 1 (module set identity).}
Each Kähler-upgrade module is declared in \texttt{src/lib.rs} under
\verb|#[cfg(feature = "kahler")]|. With the feature off, the
\texttt{rustc} configuration predicate evaluates to false at the
module declaration site, and the compiler omits the module from
the crate's item tree. Consequently the linked binary contains no
code from the omitted modules. The pre-upgrade modules are
unchanged across $E_0 \to E_i$ (verified by inspection of each
layer's diff in the engine repository), so the set of included
modules is identical.
\paragraph{Clause 2 (struct field layout).}
The only types extended by the Kähler upgrade are
\texttt{BundleSchema} and \texttt{BundleStore}, and the
extensions take the form
\verb|#[cfg(feature = "kahler")] pub <field>: <type>|. With the
feature off, the field is omitted from the struct at
\texttt{rustc} configuration time; the struct's layout is computed
from the remaining fields, which are identical to the pre-upgrade
form. Consequently the struct's field offset table, total size,
and alignment are identical between $E_0$ and $E_i$.
\paragraph{Clause 3 (test outcome identity).}
The pre-upgrade test files in \texttt{src/} and \texttt{tests/}
were not modified by any Kähler-upgrade layer commit (verified by
inspection of each layer's diff). New test files added by the
upgrade carry the \verb|#![cfg(feature = "kahler")]| attribute and
are excluded from the no-feature build's compilation unit.
Consequently the no-feature test binary built from $E_i$ contains
exactly the same test functions as the no-feature test binary
built from $E_0$, compiled against the same struct definitions
and module set per clauses 1 and 2. Pass/fail outcome is therefore
identical assuming determinism of the underlying tested
operations, which is established by the pre-upgrade engine's
existing test discipline. \qed
\end{proof}

\begin{remark}[Empirical confirmation]
\label{rem:lemma-empirical}
The lemma's clause 3 conclusion is also empirically confirmed by
Table~\ref{tab:test-counts-per-layer}: the no-feature pass count
is $720$ at every layer commit. The lemma is the structural
reason; the table is the receipt.
\end{remark}

\begin{remark}[What the lemma does not claim]
\label{rem:lemma-limits}
The lemma claims byte-equality of the no-feature build under
identical toolchain conditions. It does not claim that the
no-feature build is byte-equal across different Rust toolchain
versions, across different host platforms, or across different
linker configurations --- those introduce sources of variability
the Kähler upgrade does not control. It also does not claim that
the upgrade is the unique way to ship an additive feature; many
codebases use feature gates. What the lemma claims is that
\emph{this} upgrade, with \emph{this} discipline, satisfies the
contract.
\end{remark}

\subsection{Independent value}
\label{subsec:contract-independent}

The pattern is reusable. Any database engine maintainer shipping
a math-heavy or otherwise invasive upgrade can adopt the same
three disciplines:
\begin{enumerate}[leftmargin=*]
  \item Gate new modules behind a single compile-time feature
    flag; pin the no-feature build's test count and pass count
    in CI; reject any PR that changes them.
  \item Where existing types must grow, grow them with cfg-gated
    fields rather than new types, so the struct layout under the
    no-feature build remains byte-equal.
  \item For every consumer-facing API surface added by the
    upgrade, ship a paired contract test that exercises the
    surface's field set, variant exhaustiveness, error shape, and
    method signature; require the contract test to be in the
    same PR as the API addition.
\end{enumerate}
The discipline does not depend on the Davis framework or on any
Kähler-specific content. We tag it here as an independently
citable pattern so that database researchers shipping unrelated
math-heavy upgrades can reach for it without subscribing to the
geometry of this paper.

The Kähler upgrade itself is one specific application of the
pattern. The pattern's value is that it admitted the application
at all: without the contract, eight layers of geometric machinery
would have entailed eight rounds of consumer disruption, eight
re-validations of pre-existing behavior, and eight opportunities
to silently break a downstream client. With the contract, the
upgrade ships under a single switch, and the rest of the engine
is unchanged.
